---
title: "Chapter 5: Spending our Data"
output: html_notebook
---

```{r}
ames <- ames %>% 
  mutate(Sale_Price_log = log10(Sale_Price))
```

# Common Methods for Splitting Data

```{r}
library(tidymodels)
tidymodels_prefer()

set.seed(501)

# save the split information for an 80/20 split of the data
ames_split <- initial_split(ames, prop = 0.80)
ames_split
```

To get the datasets from the `rsplit` object, we need two more functions:

```{r}
ames_train <- training(ames_split)
ames_test <- testing(ames_split)

dim(ames_train)
```

Random sampling doesn't work as well when there are clear groups, or clusters, in the data. For this, we need stratified sampling.

```{r}
set.seed(502)
ames_split <- initial_split(ames, prop = 0.8, strata = Sale_Price_log)
ames_train <- training(ames_split)
ames_test <- testing(ames_split)
```

With time series data, we usually want to use the most recent data as the test set. For that, we use `initial_time_split()` which denotes what proportion of the first part of the data should be used as the training set.

# What About a Validation Set?

A validation set is a way to test how well the model will preform prior to the test set.

# Multi-Level Data

This type of data contains multiple rows per experimental unit. Data splitting should occur at the independent experimental unit level of data.

# Other Considerations
