---
title: "Chapter 6: Fitting Models with **parsnip**"
output: html_notebook
---

The **parsnip** package provides a fluent and standardized interface for a variety of models.

# Create a Model

For tidymodels, the approach to specifying a model is intended to be more unified:

1.  **Specify the *type* of model based on its mathematical structure.**
2.  **Specify the *engine* for fitting the model**. This is the software to use to fit the model.
3.  **When required, declare the *mode* of the model**. The mode reflects the type of prediction outcome. For numeric outcomes, the mode is *regression*. For qualitative outcomes, the mode is *classification*.

```{r}
library(tidymodels)
tidymodels_prefer()

linear_reg() %>% set_engine("lm")

linear_reg() %>%  set_engine("glmnet")

linear_reg() %>% set_engine("stan")
```

Model estimation can then be done with the `fit()` function or the `fit_xy()` function. The `translate()` function can provide details on how **parsnip** converts the user's code to the package's syntax:

```{r}
linear_reg() %>% set_engine("lm") %>% translate()

linear_reg(penalty = 1) %>% set_engine("glmnet") %>% translate()

linear_reg() %>% set_engine("stan") %>% translate()
```

Let's walk through how to predict the sale prices of houses in the Ames data as a function of only longitude and latitude:

```{r}
data(ames, package = "modeldata")

ames <- ames %>% 
  mutate(Sale_Price_log = log10(Sale_Price))

set.seed(502)
ames_split <- initial_split(ames, prop = 0.8, strata = Sale_Price_log)
ames_train <- training(ames_split)
ames_test <- testing(ames_split)

lm_model <- 
  linear_reg() %>% 
  set_engine("lm")

lm_form_fit <- 
  lm_model %>% 
  fit(Sale_Price_log ~ Longitude + Latitude, data = ames_train)

lm_xy_fit <- 
  lm_model %>% 
  fit_xy(
    x = ames_train %>% select(Longitude, Latitude),
    y = ames_train %>% pull(Sale_Price_log)
  )

lm_form_fit

lm_xy_fit
```

To see how **parsnip** maps arguments to the original names we can use the `translate()` function:

```{r}
rand_forest(trees = 1000, min_n = 5) %>% 
  set_engine("ranger") %>% 
  set_mode("regression") %>% 
  translate()
```

Modeling functions in **parsnip** separate model arguments into two categories:

-   *Main arguments* are more commonly used and tend to be available across engines

-   *Engine arguments* are either specific to a particular engine or using more rarely

# Using the Model Results

We examine the fitted model with `extract_fit_engine()`:

```{r}
lm_form_fit %>% extract_fit_engine()
```

We can also use `summary()`:

```{r}
model_res <- 
  lm_form_fit %>% 
  extract_fit_engine() %>% 
  summary()

model_res

param_est <- coef(model_res)
param_est
```

```{r}
tidy(lm_form_fit)
```

# Make Predictions

Another area where **parsnip** diverges from conventional R modeling is the format of values returned from `predict()`. For prediction, **parsnip** always conforms to the following rules:

1.  The results are always a tibble.
2.  The column names of the tibble are always predictable.
3.  There are always as many rows in the tibble as there are in the input data set.

```{r}
ames_test_small <- ames_test %>% slice(1:5)
predict(lm_form_fit, new_data = ames_test_small)
```

There rules make it easier to merge predictions with the original data:

```{r}
ames_test_small %>% 
  select(Sale_Price_log) %>% 
  bind_cols(predict(lm_form_fit, ames_test_small)) %>% 
  # add CI to the results
  bind_cols(predict(lm_form_fit, ames_test_small, type = "pred_int"))
```

# Creating Model Specifications

```{r}
parsnip_addin()
```
