---
title: "Chapter 7: A Model Workflow"
output: html_notebook
---

```{r}
library(tidymodels)
data(ames)
ames <- mutate(ames, Sale_Price = log10(Sale_Price))

set.seed(123)
ames_split <- initial_split(ames, prop = 0.80, strata = Sale_Price)
ames_train <- training(ames_split)
ames_test  <-  testing(ames_split)

lm_model <- linear_reg() %>% set_engine("lm")
```

The purpose of the workflow concept is to encapsulate the major pieces of the modeling process.

# Where Does the Model Begin and End?

There are several steps involved before we fit a model:

-   What predictors to use

-   How to deal with missing data

-   Variable transformations

After the model is fit we need to make decisions about model performance and evaluation.

All of these tasks compose a **model workflow**.

# Workflow Basics

The **workflows** package allows users to bind modeling and preprocessing objects together.

```{r}
library(tidymodels)
tidymodels_prefer()

lm_model <- 
  linear_reg() %>% 
  set_engine("lm")
```

A workflow always requires a **parsnip** model object:

```{r}
lm_workflow <- 
  workflow() %>% 
  add_model(lm_model)

lm_workflow
```

If our model were very simple, a standard R formula can be used as a preprocessor:

```{r}
lm_workflow <- 
  lm_workflow %>% 
  add_formula(Sale_Price ~ Longitude + Latitude)

lm_workflow
```

Workflows have a `fit()` method that can be used to create the model:

```{r}
lm_fit <- fit(lm_workflow, ames_train)
lm_fit
```

We can also use `predict()`:

```{r}
predict(lm_fit, ames_test %>% slice(1:3))
```

# Adding Raw Variables to the `workflow()`

```{r}
lm_workflow <- 
  lm_workflow %>% 
  remove_formula() %>% 
  add_variables(outcome = Sale_Price, predictors = c(Longitude, Latitude))

lm_workflow

fit(lm_workflow, ames_train)
```

# How Does a `workflow()` Use the Formula?

## Special Formulas and In-Line Functions

To fit a regression with random effects for subjects, we use:

```{r}
library(nlme)
library(lme4)
lmer(distance ~ Sex + (age | Subject), data = Orthodont)
```

But, the special attributes of the formula cannot be assessed with typical R methods.

The solution in **workflows** is an optional supplementary model formula that can be passed to `add_model()`. For example, using the previously mentioned `strata()` function, in the **survival** package, the `add_variables()` specification provides the bare column names and the actual formula given to the model is set within `add_model()`:

```{r}
library(censored)

parametric_model <- survival_reg()

parametric_workflow <- 
  workflow() %>% 
  # pass the data along as-is:
  add_variables(outcome = c(fustat, futime), predictors = c(age, rx)) %>% 
  add_model(parametric_model,
            # This formula is given to the model
            formula = Surv(futime, fustat) ~ age + strata(rx))

parametric_fit <- fit(parametric_workflow, data = ovarian)
parametric_fit
```

# Creating Multiple Workflows at Once

The **workflowset** package allows you to create combinations of workflow components which lets you fit multiple models.

We we wanted to focus on the different ways that house location is represented in the Ames data. We can create a set of formulas that capture these predictors:

```{r}
location <- list(
  longitude = Sale_Price ~ Longitude,
  latitude = Sale_Price ~ Latitude,
  coords = Sale_Price ~ Longitude + Latitude,
  neighborhood = Sale_Price ~ Neighborhood
)
```

These can be crossed with one or more models using the `workflow_set()` function.

```{r}
library(workflowsets)
location_models <- 
  workflow_set(preproc = location, models = list(lm = lm_model))
location_models

location_models$info[[1]]

extract_workflow(location_models, id = "coords_lm")
```

Now let's create model fits for each formula and save them in a new column called `fit`.

```{r}
location_models <- 
  location_models %>% 
  mutate(fit = map(info, ~ fit(.x$workflow[[1]], ames_train)))

location_models

location_models$fit[[1]]
```

# Evaluating the Test Set

Say we've concluded our model development and have settled on a final model. We can use `last_fit()` to fit the model to the entire training set and evaluate it with the testing set.

```{r}
final_lm_res <- last_fit(lm_workflow, ames_split)
final_lm_res

fitted_lm_workflow <- extract_workflow(final_lm_res)

collect_metrics(final_lm_res)

collect_predictions(final_lm_res) %>% slice(1:5)
```
